---
title: "Joshua Burrows Project 2"
date: "16 October 2020"
output: 
  github_document: 
    toc: TRUE
params: 
  day: "Monday"
---

```{r Setup, include = FALSE}
library(knitr)
library(rmarkdown)
library(tidyverse)
library(caret)
library(corrplot)
library(shiny)
opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
```

# Render Code 

The following code runs in a separate R script to render a different document for each day of the week. 

# Introduction

The day is `r params$day`. 

# Read in Data 

```{r Read}
bikes <- read_csv(file = "../Bike-Sharing-Dataset/hour.csv")

bikes$weekday <- as.factor(bikes$weekday)
levels(bikes$weekday) <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday") 

bikes$season <- as.factor(bikes$season)
levels(bikes$season) <- c("winter", "spring", "summer", "fall")

bikes$yr <- as.factor(bikes$yr)
levels(bikes$yr) <- c("2011", "2012")

bikes$mnth <- as.factor(bikes$mnth)
levels(bikes$mnth) <- c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

bikes$weathersit <- as.factor(bikes$weathersit)
levels(bikes$weathersit) <- c("very good", "good", "bad", "very bad")

bikes$holiday <- as.factor(bikes$holiday)
levels(bikes$holiday) <- c("no", "yes")

bikes %>% head()
```

# Data 

```{r Split}
dayData <- bikes %>% filter(weekday == params$day)
dayData %>% head()

set.seed(123)
dayIndex <- createDataPartition(dayData$cnt, p = .7, list = FALSE)
```

## EDA 

### Predictor Summaries and Plots

Summary statistics of rental count by season

```{r Season}
seasonSum <- dayData %>% group_by(season) %>% summarize(min = min(cnt), Q1 = quantile(cnt, probs = c(.25), names = FALSE), median = median(cnt), mean = mean(cnt), Q3 = quantile(cnt, probs = c(.75), names = FALSE), max = max(cnt))

seasonSum %>% kable()
```

Total rentals each year 

```{r Year}
yearSum <- dayData %>% group_by(yr) %>% summarize(totalRentals = sum(cnt))
yearSum %>% kable()
```

Summary statistics of rental count by dayDatath. It is probably not worth including *dayDatath* and *season* in the model, so I eliminated *season*. 

```{r dayDatath}
dayDatathSum <- dayData %>% group_by(mnth) %>% summarize(min = min(cnt), Q1 = quantile(cnt, probs = c(.25), names = FALSE), median = median(cnt), mean = mean(cnt), Q3 = quantile(cnt, probs = c(.75), names = FALSE), max = max(cnt))

dayDatathSum %>% kable()
```

Scatter plot of total rentals by hour of the day

```{r Hour}
avgRentals <- dayData %>% group_by(hr) %>% summarize(meanRentals = mean(cnt))

ggplot(avgRentals, aes(x = hr, y = meanRentals)) + geom_point() + labs(title = "Total Rentals by Hour", x = "Hour of the Day", y = "Total Rentals") 
```

Average rentals by holiday 

```{r Holiday}
dayData %>% group_by(holiday) %>% summarize(meanRentals = mean(cnt)) %>% kable()
```

Average rentals by working day. Working days are neither weekends nor holidays. I decided to eliminate this variable from the model because some of the days under consideration are weekends. 

```{r Workingday}
dayData %>% group_by(workingday) %>% summarize(meanRentals = mean(cnt)) %>% kable()
```

Average rentals by weather condition

```{r Weather}
dayData %>% group_by(weathersit) %>% summarize(meanRentals = mean(cnt)) %>% kable()
```

Scatter plot of average rentals and temperature

```{r Temp}
tempAvg <- dayData %>% group_by(temp) %>% summarize(avgRentals = mean(cnt))

ggplot(tempAvg, aes(x = temp, y = avgRentals)) + geom_point() + labs(title = "Average Rentals by Temperature", x = "Normalized Temperature", "Average Rentals")
```

Scatter plot of average rentals and feeling temperature. It does not make much sense to keep *temp* and *atemp*, so I eliminated *atemp* from the model. 

```{r aTemp}
atempAvg <- dayData %>% group_by(atemp) %>% summarize(avgRentals = mean(cnt))

ggplot(atempAvg, aes(x = atemp, y = avgRentals)) + geom_point() + labs(title = "Average Rentals by Temperature", x = "Normalized Feeling Temperature", "Average Rentals")
```

Scatter plot of average rentals by humidity 

```{r Hum}
humAvg <- dayData %>% group_by(hum) %>% summarize(avgRentals = mean(cnt))

ggplot(humAvg, aes(x = hum, y = avgRentals)) + geom_point() + labs(title = "Average Rentals by Humidity", x = "Normalized Humidity", y = "Average Rentals") 
```

Average rentals by windspeed

```{r Wind}
windAvg <- dayData %>% group_by(windspeed) %>% summarize(avgRentals = mean(cnt))

ggplot(windAvg, aes(x = windspeed, y = avgRentals)) + geom_point() + labs(title = "Average Rentals by Windspeed", x = "Normalized Windspeed", y = "Average Rentals")
```

### Correlation between Predictors 

Correlation plot of quantitative predictors. 

It does not make much sense to keep *temp* and *atemp*, so I eliminated *atemp* from the model. 

```{r Correlation}
corr <- dayData %>% select(temp, atemp, windspeed, hum) %>% cor()

corrplot(corr)

```

# Train Models

## Tree

```{r Tree}
set.seed(123)
trialTrainIndex <- sample(1:nrow(dayData), size = 100)
trialTrain <- dayData[trialTrainIndex,]
trialTrain

trialTestIndex <- sample(1:nrow(dayData), size = 50)
trialTest <- dayData[trialTestIndex,]
trialTest



set.seed(123)
tree <- train(cnt ~ yr + mnth + hr + holiday + weathersit + temp + hum + windspeed, 
              data = trialTrain, 
              method = "rpart", 
              trControl = trainControl(method = "LOOCV"))
tree
```

## Boosted Tree 

```{r Boosted Tree}
set.seed(123)
boostTree <- train(cnt ~ yr + mnth + hr + holiday + weathersit + temp + hum + windspeed, 
                   data = trialTrain, 
                   method = "gbm", 
                   trControl = trainControl(method = "LOOCV"), 
                   verbose = FALSE)
boostTree
```

# Test Models 

## Tree

```{r Test Tree}
treePreds <- predict(tree, trialTrain)
treeRMSE <- postResample(treePreds, trialTrain$cnt)[1]

boostPreds <- predict(boostTree, trialTrain)
boostRMSE <- postResample(boostPreds, trialTrain$cnt)[1]

modelPerformance <- data.frame(model = c("Non-Ensemble Tree", "Boosted Tree"), RMSE = c(treeRMSE, boostRMSE))

best <- modelPerformance %>% filter(RMSE == min(RMSE))

if(best$model == "Non-Ensemble Tree"){
  final <- tree
} else if(best$model == "Boosted Tree"){
  final <- boostTree
} else{
  stop("Error")
}

final
```

# Best Model

```{r}

```







