---
title: "Joshua Burrows Project 2"
date: "16 October 2020" 
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 

For this project, I analyzed data about the Capital bike sharing system. This data has been made available by [the UCI Machine Learning Library](https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset). 

My analysis makes use of the following packages: *knitr*, *rmarkdown*, *tidyverse*, *caret*, *corrplot*, *shiny*, and *rpart.plot*. 

# Models 

I built models to predict the number of bikes that will be rented for a given day of the week. Predictor variables include:  

- Year and month of the rental  
- Hour of the day  
- Weather, temperature, humidity, and windspeed  
- Whether the day is a holiday  

The analysis for each day of the week can be viewed here: 

[Sunday](Sunday.md)  
[Monday](Monday.md)  
[Tuesday](Tuesday.md)  
[Wednesday](Wednesday.md)  
[Thursday](Thursday.md)  
[Friday](Friday.md)  
[Saturday](Saturday.md)  

# Automation 

I didn't want to have to write seven different reports, so I automated a few things to speed up the process. The code that does the automation is here: 

```{r Auto, eval = FALSE}
library(tidyverse)
library(rmarkdown)

bikes <- read_csv(file = "../Bike-Sharing-Dataset/hour.csv")

days <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday") 

output_file <- paste0(days, ".md") 

params = lapply(days, FUN = function(x){list(day = x)})

reports <- tibble(output_file, params)

apply(reports, 
      MARGIN = 1, 
      FUN = function(x){render(input = "./BurrowsProject2Analysis.Rmd", 
                               output_file = x[[1]], 
                               params = x[[2]])})
```


















